### 2019-07-25 5_evaluation.tex
More space saving.

Most systems for HPC are highly scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
flat line is ideal. Lines that rise with node count indicate less than
ideal scaling. Most notably, Spark is typically
intended for use in industrial data center applications with task
granularities measured in tens of seconds. Spark uses a centralized
controller, which limits throughput, and this is visible in the figure
as the line for Spark immediately rises with node count. Keep in mind
that Spark is being evaluated here with a non-trivial dependence
pattern that is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

Implicitly parallel systems such as PaRSEC, StarPU and Regent that
rely on runtime analysis to build the DAG can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks that are not direct
dependencies of those executed on the current node. However, this DAG
trimming approach still requires some dynamic checks that scale with
the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Approaches based on compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
still retains dynamic checks leading to limited scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The number of dependencies per task has a strong influence on the
overhead incurred by each system, as shown in
Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
dependence pattern, when varying the number of dependencies per task
from 0 to 9.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a factor of $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. For this reason, we recommend evaluating the performance of
programming systems under realistic workloads, rather than with
trivial dependencies.

@@

Also of interest is the ability of asynchronous systems to
hide communication latency in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency vs task
granularity curves for varying amounts of
communication, determined by the number
of bytes produced by each task (and therefore communicated with each
task dependency).

@@

Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the CPU and GPU kernels perform different numbers of
operations as the GPU must be given more work to achieve peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
normalized so that the number of FLOPs is constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus overlaps communication and computation. We see that the
\lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities, where the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of MTEG (as the peak performance of the GPU and CPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: The cost of communicating data and
tasks to and from the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage at small task granularities of
very lightweight mechanisms such as those in MPI.

>>>>

Most systems for HPC are highly scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and
flat is ideal. Lines that rise with node count indicate less than
ideal scaling. Most notably, Spark is primarily
intended industrial data center applications with task
granularities measured in tens of seconds. Spark uses a centralized
controller, which limits throughput, and this is visible in the figure
as the line for Spark immediately rises with node count. Keep in mind
that Spark is being evaluated here with a non-trivial dependence
pattern that is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

Implicitly parallel systems such as PaRSEC, StarPU and Regent that
rely on runtime analysis to build the DAG can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks not directly
dependent on those executed by a given node. However, this DAG
trimming approach requires dynamic checks that scale with
the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The number of dependencies per task has a strong influence on
overhead, as shown in
Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
dependence pattern, when varying the number of dependencies per task
from 0 to 9.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. For this reason, we recommend evaluating the performance of
programming systems under realistic workloads, rather than with
trivial dependencies.

@@

Also of interest is the ability to
hide communication latency in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency with varying amounts of
communication, determined by the number
of bytes produced by each task (and therefore communicated with each
task dependency).

@@

Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
FLOP/s. Note that the kernels perform different numbers of
operations as the GPU requires more work to reach peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
normalized to keep FLOPs constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model in which data is copied to and from the GPU on each step. We test
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel.
\lstinline{w4} achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data dominates at small task
granularities, where the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of METG (as peak performance on CPU and GPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: The cost of sending data and
tasks to the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage at small task granularities of
very lightweight mechanisms such as those in MPI.

### 2019-07-25 6_related.tex
More space saving.

Parallel and distributed programming systems are most
commonly evaluated using proxy- or mini-apps, or
micro\-benchmarks. Mini-apps are explicitly derived from larger
applications and therefore have the advantage of bearing some
relationship to the original. This advantage typically does not hold
for microbenchmarks.

@@

Microbenchmarks are often easier to implement, but may not be
representative of real-world applications. From the PRK
suite~\cite{PRK14}, the PRK Stencil code has been reasonably widely
implemented, and a follow-up paper~\cite{PRKRuntimes16} compares MPI
variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
single, simple code involving 2 kernels (which combined fit in under 50 lines
of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} is also relatively widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. The benchmarks consist mostly of small
kernels for dense array computations. Neither the PRK nor the NAS
benchmark suite achieve the breadth of coverage, flexibility, or ease
of implementation of Task Bench. We believe the evidence for ease of
implementation is clear; Task Bench is implemented for more systems
than all previous mini-apps and microbenchmarks. Task Bench is also
not a fixed set of benchmarks, and it is even easier to use Task Bench
to generate a new benchmark to measure some aspect of system
performance.

System-specific benchmarks have been used to quantify specific aspects
of system performance, such as MPI communication or collective
latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
not generalize beyond the immediate system they measure.

>>>>

% Force this paragraph to page break at a hyphen because otherwise
% it's going to be nearly impossible to cut enough space out of this
% paper.
\let\oldbrokenpenalty\brokenpenalty
\brokenpenalty=0

Parallel and distributed programming systems are most
commonly evaluated using proxy- or mini-apps, or
microbenchmarks. Mini-apps are explicitly derived from larger
applications and therefore have the advantage of bearing some
relationship to the original. This advantage typically does not hold
for microbenchmarks.

\brokenpenalty=\oldbrokenpenalty

@@

Microbenchmarks can be easier to implement, but risk being
unrepresentative of real applications. PRK Stencil~\cite{PRK14}
contains 2 kernels (which combined fit in under 50 lines of C++)
and a halo exchange on a structured grid; an
evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
consist mostly of small kernels for dense array computations and has
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
coverage, flexibility, or ease of implementation of Task Bench. We
believe the evidence for ease of implementation is clear; Task Bench
is implemented for more systems than all previous mini-apps and
microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
it is even easier to use Task Bench to generate new benchmarks.

System-specific benchmarks quantify specific aspects
of system performance, such as MPI communication or collective
latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
not generalize beyond the immediate system they measure.
